\abstract{
  State-Machine Replication (SMR) is a basic building block of reliable distributed
  systems that allows to implement a reliable service out of unreliable
  components. Practical implementations of SMR that can be used in cloud
  computing would have a direct impact on the reliability of many computing 
  services. However, devising new practical SMR algorithms is hard, often
  justify a publication in a top conference.

  The problem is that SMR algorithms are not modular. Thus all the reasonging,
  proofs, and test need to be done again from scratch when devising a new algorithm. With the number of
  different situations that a practical SMR algorithm must face, a modular approach
  allowing the reuse of existing techniques is especially important.
  
  We present a framework that enables composing SMR algorithms.
  Crucially, with our framework, an algorithm  is  proved  composable  once  and
  for all in isolation. Our composition theorem, mechanically proved with Isabelle/HOL,
  guarantees that  the  composition of  any  number of composable  algorithms  is a
  correct SMR algorithm.

  The key novelty of our framework is its practical  applicability,  with  state
  of  the  art  optimization  and  reconfiguration  techniques  being  built-in:
  tolerating $k$ faults with  $k+1$  replicas,  executing  commutative  commands
  without synchronization, optimizing the execution of read-only  commands,  and
  checkpointing.

  To demonstrate our framework  we  present  two  abstract
  composable algorithms  modeling  concrete algorithms  based  on  two  types  of  quorum
  systems used in practice. We demonstrate that concrete algorithms can easily be
  obtained by refinement from the abstract algorithms and we provide TLA+ specifications 
  ready to be model-checked with the TLC model-checker.
}
